Submodular Optimization for Control of Prosumer Networks

### 2016-04-11 Article.tex
figure 3 allégée
modifs michel + new intro

The today centralized top-down architecture is set to evolve to more distributed and bi-directional systems. Furthermore, the emergence of renewable and stochastic distributed energy resources (DER) in the distribution networks will require more flexibility. Within these systems, it is assumed that multiple aggregation levels will be required in order to organize and optimize communication. Agents responsible for generation and load portfolio managment, i.e prosumers \cite{Rathnayaka2012}, will provide services (generation, load shedding, frequency regulation) against remuneration. Insure grid stability within this uncertain context appears as a complex task.

Several studies highlighted the need for increasing the storage capacity of the system. Some of them even consider electric vehicles as moving capacities that could be used for frequency regulation. In the case of fixed storage devices, the locations where they should be installed is an important question since it impacts the performances of the system. In a very static and top-down architecture, where loads and generators are fixed, this question is far from trivial. But in a smart grid scenario with bi-directional flows and agents that produce and consume depending on weather conditions, this becomes even more challenging.

@@

In this paper, we explore the use of storage in a network of prosumers whose power outputs are susceptible to change due to various external causes. In such a scenario with a high penetration of renewables, perturbations in the power distributions of the prosumers are very likely. If the balance between production and consumption is lost the system might desynchronize from the main frequency $\Omega$. If nothing is done to restaure synchrony, this could damage severely the generators and machines of the system. Since re-balancing the production and the consumption is not immediate, frequency regulation operators are paid to maintain synchrony in the meantime. Because charging and discharging storages is expensive (life cycles for instance), the less energy these operators use to control the frequency, the higher the profits. Furthermore, if we assume that we have a constrained budget, we are probably unable to place storage at every node in the network. That is, we must select a priori a subset of locations and stick with it afterwards whatever the perturbations. We might expect that all sets of locations are not equal in term of the controllability that they provide. The question that we try to answer in this paper is to find where to place the storages such that the average control energy required to maintain synchrony is minimized.

We propose to answer this question by considering the action of the storages as external controllable signals on the grid dynamics. We use a coupled oscillator network as a model for simplifying the power grid dynamic. Each node, representing a prosumer, is modeled as an oscillator with a phase angle and a natural frequency that depends on its production/consumption. The oscillators are interconnected such that their frequencies are impacted by the ones of their neighbors. As we will see, the phase angles dynamics can be written as a system of coupled second order differential equations \cite{Filatrella2008}. Under some conditions, the system synchronizes and all nodes oscillates at $\Omega$. However, if a perturbation occurs, the system might desynchronize from $\Omega$ and requires control to bring it back to the stable synchronized state. Since the electrical lines and storages have finite capacities, we have further constraints on the control signals. Obviously, we do not wish to overload lines while trying to restaure synchrony. Likewise, no battery can inject power when empty, or charge/discharge quicker than some maximum rate. We propose in this paper to incorporate these constraints within the model such that the control signals remain realistic.

This paper is organized as follows, section \ref{sec:control} presents some basic notions about control in networks, section \ref{sec:The_Model} introduces the oscillator model designed for simplifying the grid's dynamics. In section \ref{sec:control_and_submodularity} we propose how we select the subset of controlled nodes such that the average control energy is optimized. In section \ref{sec:constraints} we develop the constraints relative to the system and the batteries. Finally, in section \ref{sec:Results}, we show some results and validation.

@@

It was shown by Kalman in 1963 that a linear system $(A,B)$ is controllable if the $N \times NM $ controllability matrix $ C = [B,AB,A^2B,...,A^{N-1}B] $ has full rank, that is, $rank(C) = N$. Nevertheless, when studying the controllability of non trivial networks, using the Kalman's criterion is not a simple task since matrix C becomes huge. Furthermore, because $A^k\longrightarrow 0$ when $k\longrightarrow \infty $, numeric precision becomes quickly an issue when computing the rank of C. A different approach to study controllability in complex network was introduced later on by Lin \cite{1100557}. This relates the search of the minimum driver set to a maximum matching in a bipartite graph, which can be done with the Hopcroft-Karp algorithm in $ O(|E|\sqrt{|V|}) $ in the worst case. This requires that the system $(A,B)$ is a structured system, meaning that all elements in A and B are either fixed zeros or independant free parameters. In such a situation, structural control is extremely powerfull since it enables us to discuss a network's controllability relatively quickly even if we do not know the exact weights of the edges. Note however that the independance of the parameters is a key assumption. If A is the adjacency matrix of an undirected network for instance, independance is not verified (since A is symetric). In this case, one should use different tools to study controllability \cite{Yuan2014}.

On the other hand, if the system is controllable, there might be more than one sequence of control inputs $u(t)$ that could drive it from $X_0$ to $X_T$. Among all these possibilities, optimal control is devoted to find the sequence that minimizes some cost function that can depend on the sates of the system $X(t)$, the control inputs $u(t)$, and the final state $X_T$. In the following, these control inputs will be used to model the actions of storage devices in a power grid. We are therefore particularly interested in the sequence that requires the minimum amount of control energy \cite{Yan2012}. Let $ \mathcal{E} = \int_{0}^{T} \Vert u(t) \Vert^2 dt $ be the energy used for the control of the system. It can be shown \cite{Liu2015} that the control inputs that minimize $ \mathcal{E} $ can be written as :

@@

In this section, we introduce the coupled oscillators network model used to simplify the power grid dynamic, more details can be found in \cite{Filatrella2008}.

@@

\begin{equation}
\label{eq:bilan}
 P_{S,i}  =  P_{diss,i} + P_{acc,i} + P_{transmitted,i} 
\end{equation}

By substituting equations \ref{eq:diss}, \ref{eq:acc}, and \ref{eq:trans} into equation \ref{eq:bilan} and re-arranging the terms, we obtain the following non-linear coupled system of equations :

@@

\begin{equation}
 \ddot{ \theta_i } \sim \psi_i - \alpha \dot{ \theta_i } - \sum_{j\neq i} K_{ij} sin \left[ \theta_j - \theta_i \right] 
\end{equation}

Where $ \alpha = \frac{2 K_D}{I} $ is the dissipation term, $ K_{ij} = \frac{P_{ij}^{MAX}}{I \Omega} $ are the coupling strengths, $ \psi_i = \left[ \frac{P_{S,i}}{I \Omega} - \frac{K_D \Omega}{I} \right] $. By working in a rotating frame, we can simplify $ \psi_i=\frac{P_{S,i}}{I \Omega} $.

@@

Where $ A \circ B $ represents the Hadamard product between matrices A and B, and L is the Laplacian matrix of the underlying topology. Equation \ref{eq:with_simplification} is a continuous time second order linear system of $ N $ equations.

We transform this into a discrete-time (time step is $\Delta t$) first order linear system of $ 2 N + 1 $ equations by introducing the following vector : $ X = \left( \begin{array}{c} \theta \\ \dot{\theta} \\ 1 \end{array} \right)$ :

@@

Note that the transition matrix A encodes all the system parameters, topology, and power distribution.

We now arive at the main question of the present paper, given the dynamic of equation \ref{eq:final_equation}, how can we select the driver nodes such that we use low control energy on average ?

@@

Finding the set $ S_k $ of size k that maximizes a set function F is a difficult problem because the number of sets grows exponentially with the number of nodes. Therefore complete enumeration and evaluation is only a feasible solution on very small examples. Nevertheless, if the set function is submodular, a simple greedy heuristic returns a solution $ S_k^{\star} $ such that, in the worst case, $ \frac{F(S_k^{\star})}{F(S_k^{OPT})} \sim 63\%$ (where $ S_k^{OPT}$ is the optimum set of size k) \cite{Krause2014}. This heuristic starts with a set S (possibly empty) and iteratively adds the element i that exhibits the highest marginal gain : $ F(S \cup \{i\}) \geq F(S \cup \{j\})\ \forall j $.

@@

Recall that $\mathcal{E}_{min}$ (see eq. \ref{eq:energy}) depends on the initial and final states as well as on the inverse of the gramian matrix W, which only depends on A and B. W can thus be used to obtain information about average controllability in the state space. In the prosumer network that we consider, generators and loads ($\Psi$) are suceptible to change, meaning that initial ($X_0$) and final ($X_f$) states might also vary. In such a scenario, we prefer to aim at good performances on average, rather than very good performances in few specific cases and bad performances in all other situations.

@@

Theoretically, we are now able, for a given prosumer network, to compute A, B, and W. For some trajectory in the state space, we can use equation \ref{eq:ustar} to find the optimum control inputs. However, since we are considering a power grid as a network, we would like to verify that these inputs will ensure that the system physical constraints are respected in the process. We propose in the next section to take into account constraints related to line capacities and storage imperfections.

\section{Contraints}
\label{sec:constraints}

@@

The power that flows on line $ (i,j) \in E$ can be written as $ P_{i \longrightarrow j} = -P_{ij}^{MAX} \left( \theta_j(t) - \theta_i(t) \right) $. Flow constraints at time t are straightforward :

\begin{equation}
\label{eq:flow_cons}
\forall (i,j) \in E,\ \left|\ \theta_j(t) - \theta_i(t)\ \right| \leq 1
\end{equation}
If these constraints are verified for all instants t during the control phase then no line gets overloaded by the action of the control inputs.

\cite{Dorfler2013} showed that a condition for synchronisation in a coupled oscillators network is $ \Vert L^{\dagger}\omega \Vert_{\infty,E} \leq sin(\gamma) $, where L is the Laplacian matrix of the network, $ \omega $ is the vector of the natural frequencies of the oscillators, and $ \Vert x \Vert_{\infty,E} = max_{(i,j) \in E} \left| x_i - x_j \right| $. If this condition is satisfied, the oscillators synchronize at the common frequency $ \omega_{SYNC} $ with the phase lock $ \left| \theta_i - \theta_j \right| \leq    \gamma \in[0,\frac{\pi}{2}],\ \forall (i,j) \in E $. The synchronisation condition becomes in our settings :

As for power lines, we need to take into account the imperfections of the storage that we consider. In practice, we could use some detailed and complex models, but as to keep things simple, we adopt here a straightforward approach. A storage i has a maximum charge/discharge rate $r_i$, some amount of energy stored $ \Lambda_i(t) $ at time t, and a maximum capacity $ \Lambda_{i, MAX} $. We assume that all maximum rates ($r$) and all maximum capacities $\Lambda_{Max}$ are the same accross the storage equipments. We also denote by $ \Lambda(t) $ the vector of energy level at time t. Since $ u^{\star}(t) $ specifies the control inputs, the energy level dynamic of the batteries is $ \Lambda(t+\Delta t) = \Lambda(t) - u^{\star}(t) I \Omega $. Which can also be written as :

@@

Finding the driver set $S_k$ of size k using the gramian $W_{S_k}$ of the system $(A,B_{S_k})$ does not require initial and final states. Besides, $W_{S_k}$ does not depend on the modified power distribution $ \Psi $ of the nodes. We are indeed looking for k controllers that perform well on average over all possible situations. On the contrary, the constraints derived above (eq. \ref{eq:flow_cons}, \ref{synchro_constraint}, \ref{level_constraints} and \ref{rate_constraints}) are bound to a particular trajectory in the state space. We indeed check wether the power grid can sustain the controlled dynamics when transitioning from a given initial state to a target final state. As we do not know what these states could be, we could use multiple scenarios and test wether $S_k$ can control the system without violation of constraints. In this paper, we use a slightly different approach. We consider that the system is initially at equilibrium (all elements are synchronized at $\Omega$) and that control will be necessary if a perturbation (power inbalance) occurs and takes the system out of equilibrium (see figure \ref{fig:frequecies}). Depending on how much time we need to start the control phase, the state of the system (the initial state for control) might be "somewhere around" the synchronized state. In order to test wether a set $S_k$ can control the system without violation of constraints, we sample initial states within some hypersphere centered on the synchronized state and check all the constraints. If for all trajectories, the constraints are respected, then we consider that $S_k$ enables the control of the system.

@@

The purpose of selecting the drivers according to a Gramian related metric is to minimize, on average, the amount of control energy needed. Conversely, if we select the drivers randomly, we expect to need, on average, more energy to control the system. In figure \ref{fig:densities}, we compare the average control energy  $log(\mathcal{E})$ in function of $n_D = N_D/N_{nodes}$ for a set of drivers selected thanks to our algorithm with randomly selected drivers (in both cases overloading and battery limits constraints are satisfied). We draw $10^4$ scale-free topologies with 50 nodes, and random power distributions and line capacities.  For each system we select a random number of drivers $N_D \sim \mathcal{U}(1,N_{nodes}) $ and we find two driver sets of size $N_D $. One is chosen randomly and the other is found with algorithm \ref{algo_1}. For both sets, if the control is possible, we draw a random initial state $Y_i$ and a random final state $Y_f$, and we compute the control energy required for driving the system from $Y_i$ to $Y_f$. Since the algorithm is based on $ Tr[W^{-1}]$, we could plot $ Tr[W^{-1}]$ in function of $n_D$ but we prefer to validate our method by computing the actual average energy. As expected, both curves decrease as the number of drivers increases, meaning that, as the number of drivers grows, the average control energy tends to decrease, but we tend to use less energy when the drivers are selected with our algorithm. Note that this difference tends to zero when $n_D$ tends to one, because almost all nodes are then selected, yielding little flexibility for optimization.

@@

\begin{figure}
\includegraphics[scale=.55]{plot4.pdf}%./figure_3/figure_3}
\caption{$log(\mathcal{E})$ against $n_D$ for random and optimized driver sets ($ N_{nodes} = 50 $).}
\label{fig:densities}
\end{figure}

We investigate next how the topology of the grid and the physical constraints affect the minimum size of the driver set. The driver set size $n_D$ is no longer selected randomly, but minimized. We consider the simple case of an Erdös-Renyi topology with probability of connection p. We show on Figure \ref{fig:erdos_renyi} how the minimum size of the driver set evolves with p for different Gramian based metrics and for two levels of constraints :

\begin{itemize}
\item Level 1 : full control and grid constraints (see section \ref{sec:control_and_submodularity}). The system is controllable (i.e it can be moved from any point to any other point) and under the constraints (i.e it can be moved without overloading any line or breaking any battery limits).
\item Level 2 : full control only : the system can be moved from any point to any other point of the state space, without considering any overloading or battery limit constraints. 
\end{itemize}

@@

We believe that interesting work could be done by combining this model with real production and consumption data. There are indeed complex spatial and temporal correlations that impacts these distributions. Weather perturbations would therefore affect the nodes not completely at random, and affect how control should be designed.

>>>>

The today centralized top-down architecture is set to evolve to more distributed and bi-directional systems. Furthermore, the emergence of renewable and stochastic distributed energy resources (DER) in the distribution networks will require more flexibility. Within these systems, it is assumed that multiple aggregation levels will be required in order to organize and optimize communication. Agents responsible for generation and load portfolio managment, the so-called prosumers \cite{Parag2016} because they both produce and consume electricity, will provide services (generation, load shedding, frequency regulation) against remuneration \cite{Gensollen2014}. Insure grid stability within this uncertain context appears as a complex task.

Several studies highlighted the need for increasing the storage capacity of the system. Some of them even consider electric vehicles as moving capacities that could be used for frequency regulation. In the case of fixed storage devices, the locations where they should be installed is an important question since it impacts the performances of the system. In a very static and top-down architecture, where loads and generators are fixed, this question is far from trivial. But in a smart grid scenario with bi-directional flows and agents that produce and consume depending on weather conditions, this becomes even more challenging. In this direction, \cite{Gkatzikis2015} studies the placement of storage equipments within microgrids in a collaborative scheme. The authors solve an interesting optimization problem mixing placement-dimensioning of storage with utilization. Moreover, \cite{Gkatzikis2015} uses a Nash bargaining framework to determine how prosumers should share the costs and benefits of the storages.

@@

In this paper, we explore the use of storage in a network of prosumers whose power outputs are susceptible to change due to various external causes. In such a scenario with a high penetration of renewables, perturbations in the power distributions of the prosumers are very likely. Re-balancing the production and the consumption is then necessary and requires that we have the capability to control the grid's dynamic. Control of smart grid systems has been recently studied in \cite{Farraj2015}, where the authors use linear-quadratic optimal control theory in order to compute DER outputs. Control inputs rely on information collected by sensors and phasor measurement units and a communication network is used for sharing the information. The performances are then studied under practical limitations such as latency, sampling rate, or signal-to-noise ratio.

The present work differs in that we consider the control inputs as the actions of the storages on the grid dynamics. We are given a network of prosumers, and we have to find both the number and locations of the storages that should be deployed. Once this choice is made, we have to stick with it afterwards whatever the perturbations that may occur. We are thus looking for the smallest set of storages, as well as their locations, that will require a priori small control energy. Since we do not have any prior on the future perturbations, we minimize the average control energy required to move the system around the state space.

%If the balance between production and consumption is lost the system might desynchronize from the main frequency $\Omega$. If nothing is done to restaure synchrony, this could damage severely the generators and machines of the system. Since r

%, frequency regulation operators are paid to maintain synchrony in the meantime. Because charging and discharging storages is expensive (life cycles for instance), the less energy these operators use to control the frequency, the higher the profits. Furthermore, if we assume that we have a constrained budget, we are probably unable to place storage at every node in the network. That is,

%We might expect that all sets of locations are not equal in term of the controllability that they provide.

The power grid dynamic is represented by a coupled oscillator network. Each node, representing a prosumer, is modeled as an oscillator with a phase angle and a natural frequency that depends on its production/consumption. The oscillators are interconnected such that their frequencies are impacted by the ones of their neighbors. As we will see, the phase angles dynamics can be written as a system of coupled second order differential equations \cite{Filatrella2008}.

Optimal control theory can be used to compute the control inputs needed to restaure synchrony. However, there is no guarantee that these inputs will respect the physical constraints of the system. Electrical lines have indeed maximum capacities, and storages are also limited with bounded rates of charge/discharge which limit the possible inputs. We propose in this paper to incorporate these constraints in the model such that the control inputs remain realistic. Furthermore, we propose an optimization method that uses submodular set functions in order to find both the number of storages and their locations such that the average control energy required to maintain stability in this constrained system is minimized.

%Under some conditions, the system synchronizes and all nodes oscillates at $\Omega$. However, if a perturbation occurs, the system might desynchronize from $\Omega$ and requires control to bring it back to the stable synchronized state. Since the electrical lines and storages have finite capacities, we have further constraints on the control signals. Obviously, we do not wish to overload lines while trying to restaure synchrony. Likewise, no battery can inject power when empty, or charge/discharge quicker than some maximum rate. We propose in this paper to incorporate these constraints within the model such that the control signals remain realistic.

This paper is organized as follows, section \ref{sec:control} presents recent advances about control in networks, section \ref{sec:The_Model} introduces the oscillator model designed for simplifying the grid's dynamic and derives the constraints on the control inputs. In section \ref{sec:control_and_submodularity} we propose an optimization method for finding the subset of driver nodes such that the average control energy is optimized. Finally, in section \ref{sec:Results}, we show some results and validation.

@@

It was shown by Kalman in 1963 that a linear system $(A,B)$ is controllable if the controllability matrix $ C = [B,AB,A^2B,...,A^{N-1}B] $ has full rank, that is, $rank(C) = N$. Nevertheless, when studying the controllability of non trivial networks, using the Kalman's criterion is not a simple task since matrix C becomes huge. Furthermore, because $A^k\longrightarrow 0$ when $k\longrightarrow \infty $, numeric precision becomes quickly an issue when computing the rank of C. A different approach to study controllability in complex network was introduced later on by Lin \cite{1100557}. This relates the search of the minimum driver set to a maximum matching in a bipartite graph, which can be done with the Hopcroft-Karp algorithm in $ O(|E|\sqrt{|V|}) $ in the worst case. This requires that the system $(A,B)$ is a structured system, meaning that all elements in A and B are either fixed zeros or independant free parameters. In such a situation, structural control is extremely powerfull since it enables us to discuss a network's controllability relatively quickly even if we do not know the exact weights of the edges. Note however that the independance of the parameters is a key assumption. If A is the adjacency matrix of an undirected network for instance, independance is not verified (since A is symetric). In this case, one should use different tools to study controllability \cite{Yuan2014}.

The fact that a system is controllable or not is not the only criterion of interest. Indeed, if it is controllable, the amount of control energy required also appears as a key criterion. Actually, if the system is controllable, there might be more than one sequence of control inputs $u(t)$ that could drive it from $X_0$ to $X_T$. Among all these possibilities, optimal control is devoted to find the sequence that minimizes some cost function that can depend on the sates of the system $X(t)$, the control inputs $u(t)$, and the final state $X_T$. In the following, these control inputs will be used to model the actions of storage devices in a power grid. We are therefore particularly interested in the sequence that requires the minimum amount of control energy \cite{Yan2012}. Let $ \mathcal{E} = \int_{0}^{T} \Vert u(t) \Vert^2 dt $ be the energy used for the control of the system. It can be shown \cite{Liu2015} that the control inputs that minimize $ \mathcal{E} $ can be written as :

@@

\subsection{Grid dynamic without constraint}

In this section, we introduce the coupled oscillators network model used to simplify the power grid dynamic. More details can be found in \cite{Filatrella2008}.

@@

\begin{equation}
\label{eq:bilan}
 P_{S,i}  =  P_{diss,i} + P_{acc,i} + P_{transmitted,i},
\end{equation}

where $ P_{S,i} $ is the power of an ideal source or sink at node i. By substituting equations \ref{eq:diss}, \ref{eq:acc}, and \ref{eq:trans} into equation \ref{eq:bilan} and re-arranging the terms, we obtain the following non-linear coupled system of equations :

@@

\begin{equation}
 \ddot{ \theta_i } \sim \psi_i - \alpha \dot{ \theta_i } - \sum_{j\neq i} K_{ij} sin \left[ \theta_j - \theta_i \right],
\end{equation}

where $ \alpha = \frac{2 K_D}{I} $ is the dissipation term, $ K_{ij} = \frac{P_{ij}^{MAX}}{I \Omega} $ are the coupling strengths, $ \psi_i = \left[ \frac{P_{S,i}}{I \Omega} - \frac{K_D \Omega}{I} \right] $. In order not to overload the equations, we simplify the constant term $ \frac{K_D \Omega}{I} $ by working in a rotating frame such that $ \psi_i=\frac{P_{S,i}}{I \Omega} $.

@@

Where $ A \circ B $ represents the Hadamard product between matrices A and B, and L is the Laplacian matrix of the underlying topology ($L_{ij} = k_i $ if $i=j$ and $L_{ij} = -m_{ij} $ otherwise). Equation \ref{eq:with_simplification} is a continuous time second order linear system of $ N $ equations.

By introducing $ X = \left( \begin{array}{c} \theta \\ \dot{\theta} \\ 1 \end{array} \right)$, we transform this into a first order linear system of $2N+1$ equations, which is discretized with time step $\Delta t$, leading to :

@@

Note that the transition matrix A encodes all the system parameters, topology, and power distribution. So far we have expressed the grid dynamic as a coupled oscillators network, but we did not incorporate the different physical constraints on the network.

\cite{Dorfler2013} showed that a condition for synchronisation in a coupled oscillators network is $ \Vert L^{\dagger}\omega \Vert_{\infty,E} \leq sin(\gamma) $, where $ L^{\dagger}$ is the Moore-Penrose pseudo inverse of the Laplacian matrix of the network, $ \omega $ is the vector of the natural frequencies of the oscillators, and $ \Vert x \Vert_{\infty,E} = max_{(i,j) \in E} \left| x_i - x_j \right| $. If this condition is satisfied, the oscillators synchronize at the common frequency $ \omega_{SYNC} $ with the phase lock $ \left| \theta_i - \theta_j \right| \leq    \gamma \in[0,\frac{\pi}{2}],\ \forall (i,j) \in E $.

On the other hand, the power that flows on line $ (i,j) \in E$ can be written as $ P_{i \longrightarrow j} = -P_{ij}^{MAX} sin\left( \theta_j(t) - \theta_i(t) \right) $. If $ \theta_j(t) - \theta_i(t) \sim 0 $, the flow constraints at time t are straightforward :

\begin{equation}
\label{eq:flow_cons}
\forall (i,j) \in E,\ \left|\ \theta_j(t) - \theta_i(t)\ \right| \leq 1
\end{equation}
If these constraints are verified for all instants t during the control phase then no line gets overloaded by the action of the control inputs. Writing the synchronisation condition in our settings thus gives :

A storage i has a maximum charge/discharge rate $r_i$, some amount of energy stored $ \Lambda_i(t) $ at time t, and a maximum capacity $ \Lambda_{i, MAX} $. We assume that all maximum rates ($r$) and all maximum capacities $\Lambda_{Max}$ are the same accross the storage equipments. We also denote by $ \Lambda(t) $ the vector of energy level at time t. Since $ u^{\star}(t) $ specifies the control inputs, the energy level dynamic of the batteries is $ \Lambda(t+\Delta t) = \Lambda(t) - u^{\star}(t) I \Omega $. Which can also be written as :

We now arrive at the main question of the present paper, given the dynamic of equation \ref{eq:final_equation} and the constraints of equations \ref{eq:flow_cons}, \ref{synchro_constraint}, \ref{zero_sum_constraint}, \ref{level_constraints}, and \ref{rate_constraints}, how can we select the driver nodes such that we use low control energy on average ?

@@

Generally speaking, finding the set $ S_k $ of size k that maximizes a set function F is a difficult problem because the number of sets grows exponentially with the number of nodes. Therefore complete enumeration and evaluation is only a feasible solution on very small examples. Nevertheless, if the set function is submodular, a simple greedy heuristic returns a solution $ S_k^{\star} $ such that, in the worst case, $ \frac{F(S_k^{\star})}{F(S_k^{OPT})} \sim 63\%$ (where $ S_k^{OPT}$ is the optimum set of size k) \cite{Krause2014}. This heuristic starts with a set S (possibly empty) and iteratively adds the element i that exhibits the highest marginal gain : $ F(S \cup \{i\}) \geq F(S \cup \{j\})\ \forall j $.

@@

Recall that $\mathcal{E}_{min}$ (see eq. \ref{eq:energy}) depends on the initial and final states as well as on the inverse of the gramian matrix W, which only depends on A and B. W can thus be used to obtain information about the average control energy required to move the system in the state space. In the prosumer network that we consider, generators and loads ($\Psi$) are suceptible to change, meaning that initial ($X_0$) and final ($X_f$) states might also vary. In such a scenario, we prefer to aim at good performances on average, rather than very good performances in few specific cases and bad performances in all other situations.

@@

Theoretically, we are now able, for a given prosumer network, to compute A, B, and W. For some trajectory in the state space, we can use equation \ref{eq:ustar} to find the optimum control inputs. However, these inputs do not take into account the physical constraints on the possible trajectories due to lines and batteries finite capacities.

@@

@@

Finding the driver set $S_k$ of size k using the gramian $W_{S_k}$ of the system $(A,B_{S_k})$ does not require initial and final states. Besides, $W_{S_k}$ does not depend on the modified power distribution $ \Psi $ of the nodes. We are indeed looking for k controllers that perform well on average over all possible situations. On the contrary, the constraints derived above (eq. \ref{eq:flow_cons}, \ref{synchro_constraint}, \ref{level_constraints} and \ref{rate_constraints}) are bounded to a particular trajectory in the state space. We indeed check wether the power grid can sustain the controlled dynamics when transitioning from a given initial state to a target final state. As we do not know what these states could be, we could use multiple scenarios and test wether $S_k$ can control the system without violation of constraints. In this paper, we use a slightly different approach. We consider that the system is initially at equilibrium (all elements are synchronized at $\Omega$) and that control will be necessary if a perturbation (power inbalance) occurs and takes the system out of equilibrium (see figure \ref{fig:frequecies}). Depending on how much time we need to start the control phase, the state of the system (the initial state for control) might be "somewhere around" the synchronized state. In order to test wether a set $S_k$ can control the system without violation of constraints, we sample initial states within some hypersphere centered on the synchronized state and check all the constraints. If for all trajectories, the constraints are respected, then we consider that $S_k$ enables the control of the system.

@@

The purpose of selecting the drivers according to a Gramian related metric is to minimize, on average, the amount of control energy needed. Conversely, if we select the drivers randomly, we expect to need, on average, more energy to control the system. In figure \ref{fig:densities}, we compare the average control energy  $log(\mathcal{E})$ in function of the proportion of drivers amoung the total number of nodes $n_D = N_D/N_{nodes}$, for a set of drivers selected thanks to our algorithm with randomly selected drivers (in both cases overloading and battery limits constraints are satisfied). We draw $10^4$ scale-free topologies with 50 nodes, and random power distributions and line capacities.  For each system we select a random number of drivers $N_D \sim \mathcal{U}(1,N_{nodes}) $ and we find two driver sets of size $N_D $. One is chosen randomly and the other is found with algorithm \ref{algo_1}. For both sets, if the control is possible, we draw a random initial state $Y_i$ and a random final state $Y_f$, and we compute the control energy required for driving the system from $Y_i$ to $Y_f$. Since the algorithm is based on $ Tr[W^{-1}]$, we could plot $ Tr[W^{-1}]$ in function of $n_D$ but we prefer to validate our method by computing the actual average energy. As expected, both curves decrease as the number of drivers increases, meaning that, as the number of drivers grows, the average control energy tends to decrease, but we tend to use less energy when the drivers are selected with our algorithm. Note that this difference tends to zero when $n_D$ tends to one, because almost all nodes are then selected, yielding little flexibility for optimization.

@@

\begin{figure}
\includegraphics[scale=.55]{plot4.png}%./figure_3/figure_3}
\caption{$log(\mathcal{E})$ against $n_D$ for random and optimized driver sets ($ N_{nodes} = 50 $).}
\label{fig:densities}
\end{figure}

We investigate next how the topology of the grid and the physical constraints affect the minimum size of the driver set. The driver set size $n_D$ is no longer selected randomly, but minimized. We consider the simple case of an Erdös-Rényi topology with probability of connection p. We show on Figure \ref{fig:erdos_renyi} how the minimum size of the driver set evolves with p for different Gramian based metrics and for two levels of constraints :

\begin{itemize}
\item Level 1 : full control and grid constraints (see section \ref{sec:The_Model}). The system is controllable (i.e it can be moved from any point to any other point) and under the constraints (i.e it can be moved without overloading any line or breaking any battery limits).
\item Level 2 : full control only : the system can be moved from any point to any other point of the state space, without considering any overloading or battery limit constraints. 
\end{itemize}

@@

We believe that interesting work could be done by combining this model with real production and consumption data. There are indeed complex spatial and temporal correlations that impacts these distributions \cite{Gensollen2014}. Weather perturbations would therefore affect the nodes not completely at random, and affect how control should be designed.
